### 2014-04-01 paper.tex
Section about fusion for Pull arrays

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}
\usepackage{longtable}
\usepackage{stmaryrd}
\usepackage[mathletters]{ucs}
\usepackage[utf8x]{inputenc}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
\copyrightyear{20yy} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\doi{nnnnnnn.nnnnnnn}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Title Text}
\subtitle{Subtitle Text, if any}

\authorinfo{Name1}
           {Affiliation1}
           {Email1}
\authorinfo{Lenore M. Mullin}
           {Emeritus Professor\\Department of Computer Science \\
University at Albany, SUNY}
           {lenore@cs.albany.edu}

\maketitle

\begin{abstract}
This is the text of the abstract.
\end{abstract}

\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore, 
% you may leave them out
\terms
term1, term2

\keywords
keyword1, keyword2

\section{Introduction}

\section{Overview of MoA}
\input{intromoa}

\section{Overview of Pull arrays}

The functional programming community has recently gained an increased
interest in high performance parallel array programming
\cite{keller2010regular,Axelsson:2010:Feldspar,Mainland:2010:Nikola,Svensson:2011:Obsidian,Claessen:2012:Expressive,Ankner:2013:AnEDSL,lippmeier2011efficient}. One
of the central abstractions in this line of work is the \emph{pull
  array}, also know as delayed array. The can be defined as follows
(we use Haskell \cite{marlow2010haskell} to demonstrate functions
programs).

\begin{verbatim}
data Pull a = Pull { ixf    :: (Int -> a)
                   , length :: Int
                   }
\end{verbatim}

Arrays are represented as functions from index to element. We refer to
this function as an \emph{index function}. Additionally arrays also
has a length.

This representation has several advantages:
\begin{itemize}
\item \emph{Parallelism}. Since each element is computed
  independently, they can easily be computed in parallel.
\item \emph{Fusion}. When functions on pull arrays are composed, the intermediate
  arrays can be automatically removed. 
\item \emph{Compositionality}. Pull arrays provide a wealth of
  compositional and highly reusable combinators. These can be composed
  together to write high-level programs which typically are easier to
  understand than monolithic array processing code.
\end{itemize}

\begin{figure}
\begin{verbatim}
map :: (a -> b) -> Pull a -> Pull b
map f (Pull ixf l) = Pull (f . ixf) l
\end{verbatim}
\caption{Example functions on pull arrays}
\end{figure}

\subsection{Fusion}

As mentioned, one of the most important advantages of Pull arrays is
that they support
fusion\cite{gill1993short,axelsson2010feldspar,keller2010regular}. Fusion
is a program transformation which removes intermediate data structures
and is a decendant of deforestation
\cite{wadler1990deforestation}. Fusion supercedes transformations such
as loop fusion in optimizing compilers for imperative languages.  Pull
arrays make it particularly simple to achieve fusion. It simply
amounts to inlining the definitions of functions.

As an example of fusion consider the example
\verb!map (*2) (map (*5) arr)!, where each element of the array
\verb!arr! is first multiplied by five and then again by
two. Conceptually, there is an intermediate array created after the
first \verb!map! but thanks to fusion it will be removed. Fusion will proceed
as follows:

\begin{verbatim}
map (*2) (map (*5) arr)
=> { arr }
map (*2) (map (*5) (Pull ixf l)
=> { map }
map (*2) (Pull ((*5) . ixf) l)
=> { map }
Pull ((*2) . (*5) . ixf) l)
\end{verbatim}

The result is a single Pull array, the intermediate array has been removed.

The correctness of fusion relies crucially on the language being
\emph{pure}. If side effects where allowed in the indexing function
fusion would not be correct.

There are two different ways of implementing Pull arrays in the
literature. The simplest one comes from using Pull arrays in an
embedded langauge, i.e. when the array language is implemented as a
library in another host language. Examples of such languages are
Pan\cite{elliott2003compiling}, Feldspar\cite{Axelsson:2010:Feldspar},
Obsidian\cite{Svensson:2011:Obsidian} and
Nikola\cite{Mainland:2010:Nikola}. By using the technique of
implementing Pull arrays as \emph{shallow embeddings} on top of a
\emph{deeply embedded} code language, fusion comes completely for free
\cite{svenningsson2013combining}.

In the array library Repa\cite{keller2010regular}, compiler supported
rewrite rules are used to implement fusion \cite{jones2001playing}.

\subsection{Higher dimensions}

\section{Comparison}

\subsection{Types}

\subsection{\(\psi\) reductions and fusion}

\subsection{Higher dimensions}

\subsection{Functions}

\begin{tabular}{|@{}l|l@{}|}
\hline
MoA & Feldspar
\\
\hline
δ & dim . extent
\\
ρ & extent
\\
ιξⁿ & -- generalized enumFromTo
\\
ψ & index
\\
rav & -- flattens to a one-dimensional array
\\
γ & toIndex
\\
γ' & fromIndex
\\
$s \: \hat{ρ} \: ξ$ & reshape
\\
π x & product
\\
τ & size
\\
++ & ++
\\
ξ₁ f ξ₂ & zipWith f ξ₁ ξ₁
\\
σ f ξ & map f ξ
\\
ξ f σ & map f ξ
\\
$\bigtriangleup$ & take
\\
$\bigtriangledown$ & drop
\\
$op^{red}$ & fold
\\
Φ & reverse
\\
Θ & rotate
\\
$\varobslash$ & -- generalized transpose
\\
$f Ω_d ξ$ & -- Apply $f$ along the dimension $d$ of $ξ$
\\
\hline
\end{tabular}

\subsection{Example programs}

\section{Related work}

\section{Future work}

\section{Conclusion}

\acks

Acknowledgments.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}
\bibliography{bibliography,hpf,paper1,workshop}

% The bibliography should be embedded for final submission.

%% \begin{thebibliography}{}
%% \softraggedright

%% \bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
%% P. Q. Smith, and X. Y. Jones. ...reference text...

%% \end{thebibliography}

\end{document}